---
title: "Simple toolkit template"
echo: true
---

## Getting started

There are many options for each of the steps here, this is a minimal, quick example designed to get you something that works quickly and can be modified. *In particular, we would expect the aggregation sequences and figures to need to be tailored for each analysis.*

See (in development) [documentation website](https://mdbauth.github.io/WERP_toolkit_demo/) for more detailed examples and discussion of capabilities.

Each step saves metadata about its parameters, which can be used for parameter-based runs to replicate, as well as variables to use in plots.

```{r setup}
#| include: false 
#| #| warning: false 
#| #| message: false
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file(), cache.lazy = FALSE, warning = FALSE, message = FALSE)

```

### Load libraries

```{r}
#| message: false
library(HydroBOT)
library(sf)
library(dplyr)
library(ggplot2)
library(tidyverse)
#devtools::install_github("galenholt/hydrogauge")
library(hydrogauge)
```


## User arguments

As a user, we need to set

-   Paths to input

-   Paths to save output

-   Module arguments (minimal)

-   Aggregation sequence and functions

-   Figures (typically interactive)

## Paths

```{r}
# Outer directory for project data
project_dir = getwd()

# Hydrographs (expected to exist already)
hydro_dir = file.path(project_dir, 'data', 'hydrographs')

# Generated data
EWR_subdir <- 'LowerDarling'

# EWR outputs (created here automatically in controller, need to read from here to aggregate)
ewr_results <- file.path(project_dir, 'data', 'module_output', 'EWR', EWR_subdir)

# outputs of aggregator. There may be multiple modules
agg_results <- file.path(project_dir, 'data', 'aggregator_output')
```

## FUNCTIONS
```{r}
make_ewrformat <- function(x) {
  x <- x |> 
    dplyr::mutate(gauge = dplyr::case_when(
      grepl('ml/d', units, ignore.case = TRUE) ~ paste0(gauge, '_flow'),
      units %in% c('m', 'Res. Level AHD', 'Metres') ~ paste0(gauge, '_level'),
      .default = paste0(gauge, '_FAIL'))
      ) |> 
    dplyr::select(Date = time, gauge, value) |> 
    tidyr::pivot_wider(id_cols = Date, names_from = gauge, values_from = value) |> 
    dplyr::mutate(Date = as.Date(Date))
  return(x)
}

# bom_simple <- purrr::map(bom_ts, make_ewrformat)
# state_simple <- purrr::map(state_ts, make_ewrformat)

```

## unsing Hydrogauge
# Gauge puller
```{r}
LD_gauges <- c("425013", "425020", "425007", "425012", "425022", "425023", "425010", "414203",
               "A4260509", "A4260505", "A4260507", "425019")

#NSW gauges
ds_NSW <- get_datasources_by_site(portal = 'NSW', 
                              site_list = LD_gauges)

var_info_NSW <- get_variable_list(portal = 'NSW', 
                              site_list = LD_gauges, 
                              datasource = "A")

ts_days_NSW <- get_ts_traces(portal = 'NSW', 
                         site_list = LD_gauges[!LD_gauges %in% c("A4260509", "A4260505",
                                                                 "A4260507")], #414203 better pulled from vic
                         datasource = "A", 
                         var_list = c("100", "130", "141"),
                         start_time = 20200101,
                         end_time = 20250408,
                         interval = "day",
                         data_type = "mean", #is mean the best to use? could use 'tot' (sums)
                         multiplier = 1,
                         returnformat = 'df')

ts_days_NSW |> 
  filter(!variable %in% c("100.10", "100.09"))|>
  ggplot(aes(x = time, y = value, color = variable_short_name)) +
  geom_line() +
  facet_grid(variable_short_name ~ site_short_name, scales = 'free')

#VIC gauges
set2 <- LD_gauges[!LD_gauges %in% unique(ts_days_NSW$site)]

var_info_VIC <- get_variable_list(portal = 'Vic', 
                              site_list = LD_gauges, 
                              datasource = "A")

ts_days_VIC <- get_ts_traces(portal = 'Vic', 
                         site_list = set2,
                         datasource = "A", 
                         var_list = c("100", "130", "141.50"),
                         start_time = 20200101,
                         end_time = 20250408,
                         interval = "day",
                         data_type = "mean", #is mean the best to use? could use 'tot' (sums)
                         multiplier = 1,
                         returnformat = 'df')

ts_days_VIC |> 
  filter(!variable %in% c("100.10", "100.09"))|>
  ggplot(aes(x = time, y = value, color = variable_short_name)) +
  geom_line() +
  facet_grid(variable_short_name ~ site_short_name, scales = 'free')

#SA gauges
station_tslist <- getTimeseriesList(portal = 'bom', station_no = c("A4260509", "A4260505", "A4260507"))

station_tslist |> 
  dplyr::filter(grepl('DMQaQc.Merged.DailyMean.24HR', ts_name)) |> 
  tidyr::pivot_longer(cols = c(from, to), names_to = 'startend', values_to = 'date') |> 
ggplot(aes(y = date, x = station_no, color = parametertype_name)) +
  geom_point(position = position_dodge(width = 0.5)) + 
  geom_line(position = position_dodge(width = 0.5)) +
  coord_flip()

RM_ts <- getTimeseriesList(portal = 'bom',
                           extra_list = list(station_name = 'River Murray*',
                                             ts_name = 'DMQaQc.Merged.DailyMean.24HR'),
                           returnfields = c('station_no', 'station_name',
                                            'ts_name', 'ts_id', 
                                            'ts_unitname', 'parametertype_name',
                                            'coverage'))

RM_ts <- RM_ts |>
  filter(station_no %in% c("A4260509", "A4260505", "A4260507"))


ts_days_SA <- getTimeseriesValues(portal = 'bom',
                                  ts_id = unique(RM_ts$ts_id),
                                  start_time = 20200101,
                                  end_time = 20250408)

ts_days_SA |> 
  ggplot(aes(x = time, y = value, color = parametertype_name)) +
  geom_line() +
  facet_grid(parametertype_name ~ station_no, scales = 'free', labeller = label_wrap_gen(10))


```

# tidy data for EWR tool
```{r}
SA_simple <- purrr::map(ts_days_SA, make_ewrformat)
NSW_simple <- purrr::map(ts_days_NSW, make_ewrformat)
VIC_simple <- purrr::map(ts_days_NSW, make_ewrformat)

#Wide format for EWR tool
hydrographs_LowerDarling <- bind_cols(SA_simple, NSW_simple, VIC_simple) 

write.csv(hydrographs_LowerDarling, file.path(project_dir, "hydrographs_LowerDarling", paste0("hydrographs_LowerDarling_20250408", ".csv")), row.names = FALSE)

```



```{r}
# the way i did it before galen gave me his function.
ts_days_NSW <- ts_days_NSW |> 
  mutate(Date = format(as.POSIXct(time, format="%Y-%m-%d %H:%M:%S"), format= "%Y-%m-%d"))|>
  select(-time)

ts_days_VIC <- ts_days_VIC |> 
  mutate(Date = format(as.POSIXct(time, format="%Y-%m-%d %H:%M:%S"), format= "%Y-%m-%d"))|>
  select(-time)

ts_days_SA <- ts_days_SA |>
  mutate(site = station_no,
         variable_short_name = parametertype_name,
         Date = format(as.POSIXct(time, format="%Y-%m-%d %H:%M:%S"), format= "%Y-%m-%d"))|>
  select(-time)

# all variables
ts_days <- bind_rows(ts_days_NSW, ts_days_VIC, ts_days_SA)

# ts_days |>
#   ggplot(aes(x = Date, y = value, color = variable_short_name)) +
#   geom_line() +
#   facet_grid(variable_short_name ~ site, scales = 'free')

ts_days |>
  dplyr::distinct(site, variable)

hydrographs_LowerDarling_days <- ts_days |>
  filter(!variable %in% c("100.10", "100.09", "100.01"))|>
  filter(variable_short_name != "Electrical Conductivity @ 25C")|>
  mutate(gauge = case_when(variable_short_name %in% c("Level     (Metres)", "Res. Level AHD",
                                                      "Water Level (m)", "Water Course Level"
                                                      ) ~ paste0(site, "_level"),
                           variable_short_name %in% c("Discharge (ML/d)", "Water Course Discharge"
                                                      ) ~ paste0(site, "_flow")))

hydrographs_LowerDarling <- hydrographs_LowerDarling_days |>
  select(Date, value, gauge) |>
  pivot_wider(names_from = gauge, values_from = value)

hydrographs_LowerDarling_days|>
  ggplot(aes(x = Date, y = value, color = gauge)) +
  geom_point() +
  facet_wrap(gauge~., scales = 'free_y')

# write.csv(hydrographs_LowerDarling, file.path(project_dir, "hydrographs_LowerDarling", paste0("hydrographs_LowerDarling_20250408", ".csv")), row.names = FALSE)

```

## Plot Hydrographs:
```{r}

```


