---
title: "Simple toolkit template"
echo: true
---

## Getting started

There are many options for each of the steps here, this is a minimal, quick example designed to get you something that works quickly and can be modified. *In particular, we would expect the aggregation sequences and figures to need to be tailored for each analysis.*

See (in development) [documentation website](https://mdbauth.github.io/WERP_toolkit_demo/) for more detailed examples and discussion of capabilities.

Each step saves metadata about its parameters, which can be used for parameter-based runs to replicate, as well as variables to use in plots.

```{r setup}
#| include: false 
#| #| warning: false 
#| #| message: false
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file(), cache.lazy = FALSE, warning = FALSE, message = FALSE)

```

### Load libraries

```{r}
#| message: false
library(HydroBOT)
library(sf)
library(dplyr)
library(ggplot2)
library(tidyverse)
#devtools::install_github("galenholt/hydrogauge")
library(hydrogauge)
```


## User arguments

As a user, we need to set

-   Paths to input

-   Paths to save output

-   Module arguments (minimal)

-   Aggregation sequence and functions

-   Figures (typically interactive)

## Paths

```{r}
# Outer directory for project data
project_dir = getwd()

# Hydrographs (expected to exist already)
hydro_dir = file.path(project_dir, 'data', 'hydrographs')

# Generated data
EWR_subdir <- 'LowerDarling'

# EWR outputs (created here automatically in controller, need to read from here to aggregate)
ewr_results <- file.path(project_dir, 'data', 'module_output', 'EWR', EWR_subdir)

# outputs of aggregator. There may be multiple modules
agg_results <- file.path(project_dir, 'data', 'aggregator_output')
```

## FUNCTIONS
```{r}
make_ewrformat <- function(x) {
  x <- x |> 
    dplyr::mutate(gauge = dplyr::case_when(
      grepl('ml/d', units, ignore.case = TRUE) ~ paste0(gauge, '_flow'),
      units %in% c('m', 'Res. Level AHD', 'Metres') ~ paste0(gauge, '_level'),
      .default = paste0(gauge, '_FAIL'))
      ) |> 
    dplyr::select(Date = time, gauge, value) |> 
    tidyr::pivot_wider(id_cols = Date, names_from = gauge, values_from = value) |> 
    dplyr::mutate(Date = as.Date(Date))
  return(x)
}

# bom_simple <- purrr::map(bom_ts, make_ewrformat)
# state_simple <- purrr::map(state_ts, make_ewrformat)

```

## unsing Hydrogauge
# Gauge puller
```{r}
LD_gauges <- c("425013", "425020", "425007", "425012", "425022", "425023", "425010", "414203",
               "A4260509", "A4260505", "A4260507", "425019")

NSW_gauges <- c("425013", "425020", "425007", "425012", "425022", "425023", "425010", "425019")

#NSW gauges
ds_NSW <- get_datasources_by_site(portal = 'NSW', 
                              site_list = NSW_gauges)

var_info_NSW <- get_variable_list(portal = 'NSW', 
                              site_list = NSW_gauges, 
                              datasource = "A")

#LEVEL GAUGES:
var_info_NSW_level <- var_info_NSW |>
  filter(variable %in% c('100.00', '130.00'))

ts_days_NSW_level <- data.frame()

for (i in 1:nrow(var_info_NSW_level)) {

ts_days_NSW_temp <- get_ts_traces(portal = 'NSW', 
                         site_list = var_info_NSW_level$site[i],
                         datasource = "A", 
                         var_list = var_info_NSW_level$variable[i],
                         start_time =  as.numeric(format(as.Date(var_info_NSW_level$period_start[i]), format = "%Y%m%d")),
                         end_time = as.numeric(format(as.Date(var_info_NSW_level$period_end[i]), format = "%Y%m%d")),
                         interval = "day",
                         data_type = "mean", #is mean the best to use? could use 'tot' (sums)
                         multiplier = 1,
                         returnformat = 'df')

ts_days_NSW_level <- bind_rows(ts_days_NSW_level, ts_days_NSW_temp)

}

#FLOW GAUGES:
var_info_NSW_flow <- var_info_NSW |>
  filter(variable %in% c('151.00'))

ts_days_NSW_flow <- data.frame()

for (i in 1:nrow(var_info_NSW_flow)) {

ts_days_NSW_temp <- get_ts_traces(portal = 'NSW', 
                         site_list = var_info_NSW_flow$site[i],
                         datasource = "A", 
                         var_list = var_info_NSW_flow$variable[i],
                         start_time =  as.numeric(format(as.Date(var_info_NSW_flow$period_start[i]), format = "%Y%m%d")),
                         end_time = as.numeric(format(as.Date(var_info_NSW_flow$period_end[i]), format = "%Y%m%d")),
                         interval = "day",
                         data_type = "mean", #doesnt matter for level for 151.00
                         multiplier = 1,
                         returnformat = 'df')

ts_days_NSW_flow <- bind_rows(ts_days_NSW_flow, ts_days_NSW_temp)

}

ts_days_NSW_flow$gauge_type <-  "flow"
ts_days_NSW_level$gauge_type <- "level"


ts_days_NSW <- bind_rows(ts_days_NSW_flow, ts_days_NSW_level) 

ts_days_NSW|> 
  #filter(!variable %in% c("100.10", "100.09"))|>
  ggplot(aes(x = time, y = value, color = variable_short_name)) +
  geom_line() +
  facet_grid(site ~ data_type, scales = 'free')


```


```{r}
#VIC gauges
VIC_gauges <-  c("414203", "414209")

var_info_VIC <- get_variable_list(portal = 'Vic', 
                              site_list = VIC_gauges, 
                              datasource = "A")

var_info_VIC_level <- var_info_VIC |>
  filter(variable %in% c('100.00', '400'))

ts_days_VIC <- get_ts_traces(portal = 'Vic', 
                         site_list = VIC_gauges,
                         datasource = "A", 
                         var_list = c("100"),
                         start_time = as.numeric(format(as.Date(min(var_info_VIC_level$period_start, na.rm = TRUE)), format = "%Y%m%d")),
                         end_time = as.numeric(format(as.Date(max(var_info_VIC_level$period_end, na.rm = TRUE)), format = "%Y%m%d")),
                         interval = "day",
                         data_type = "mean", #is mean the best to use? could use 'tot' (sums)
                         multiplier = 1,
                         returnformat = 'df')

ts_days_VIC$gauge_type <- "level"

ts_days_VIC |> 
  ggplot(aes(x = time, y = value, color = variable_short_name)) +
  geom_line() +
  facet_grid(variable_short_name ~ site, scales = 'free')

```


```{r}
#SA gauges
station_tslist <- getTimeseriesList(portal = 'bom', station_no = c("A4260509", "A4260505", "A4260507",
                                                                   "A4260506", "A4260501", "A4260508")) 
                                                                    # also need last 3 gauges for some EWRs.

RM_ts <- getTimeseriesList(portal = 'bom',
                           extra_list = list(station_name = 'River Murray*',
                                             ts_name = 'DMQaQc.Merged.DailyMean.24HR'),
                           returnfields = c('station_no', 'station_name',
                                            'ts_name', 'ts_id', 
                                            'ts_unitname', 'parametertype_name',
                                            'coverage'))

RM_ts1 <- RM_ts |>
  filter(station_no %in% c(#"A4260509", "A4260505", "A4260507",
                           "A4260506", "A4260501", "A4260508"))

ts_days_SA1 <- getTimeseriesValues(portal = 'bom',
                                  ts_id = unique(RM_ts1$ts_id),
                                  #start_time = 20200101,
                                  #end_time = 20250408,
                                  period = 'complete')

RM_ts2 <- RM_ts |>
  filter(station_no %in% c("A4260509", "A4260505", "A4260507"))#,
                           #"A4260506", "A4260501", "A4260508"))

ts_days_SA2 <- getTimeseriesValues(portal = 'bom',
                                  ts_id = unique(RM_ts2$ts_id),
                                  #start_time = 20200101,
                                  #end_time = 20250408,
                                  period = 'complete')

ts_days_SA <- bind_rows(ts_days_SA1, ts_days_SA2)

ts_days_SA <- ts_days_SA |>
  filter(parametertype_name %in% c("Water Course Level", "Water Course Discharge", "Storage Level"))|>
  mutate(gauge_type = case_when(parametertype_name %in% c("Water Course Level", "Storage Level") ~ "level",
                           parametertype_name %in% c("Water Course Discharge") ~ "flow"),
         site = station_no)

#"A4260509", "A4260505", "A4260507",
#"A4260506", "A4260501", "A4260508"
ts_days_SA |> 
  ggplot(aes(x = time, y = value, color = parametertype_name)) +
  geom_line() +
  facet_grid(parametertype_name ~ station_no, scales = 'free', labeller = label_wrap_gen(10))


```

# tidy data for EWR tool
```{r}
#Manually add euston - 414203
X414203_colnames <-  colnames(read.csv(file.path(project_dir, "data", "old_hydrographs", "414203_historical.csv"),skip = 2))
X414203_historical <-  read.csv(file.path(project_dir, "data", "old_hydrographs", "414203_historical.csv"),
                                skip = 4)
colnames(X414203_historical) <- X414203_colnames

X414203_historical <- X414203_historical |>
  mutate(time = as.POSIXct(Date, format = "%H:%M:%S %d/%m/%Y")) |> #,
                       #"%Y/%m/%d %H:%M:%S")) |>
  select(time, level = Water.Level..m., flow = Discharge..ML.Day.)|>
  pivot_longer(cols = c(level, flow), names_to = "gauge_type", values_to =  "value", values_drop_na = TRUE)|>
  mutate(site = "414203")

#Manually add euston - 414209
X414209_colnames <-  colnames(read.csv(file.path(project_dir, "data", "old_hydrographs", "414209_historical.csv"),skip = 2))

X414209_historical <-  read.csv(file.path(project_dir, "data", "old_hydrographs", "414209_historical.csv"),
                                skip = 4)
colnames(X414209_historical) <- X414209_colnames

X414209_historical <- X414209_historical |>
  mutate(time = as.POSIXct(Date, format = "%H:%M:%S %d/%m/%Y")) |> #,
                       #"%Y/%m/%d %H:%M:%S")) |>
  select(time, level = Water.Level..m.)|>
  pivot_longer(cols = c(level), names_to = "gauge_type", values_to =  "value", values_drop_na = TRUE)|>
  mutate(site = "414209")

ts_days_VIC <- bind_rows(X414203_historical, X414209_historical)

```


```{r}
#long format for plotting
hydrographs_LowerDarling_long <- bind_rows(ts_days_NSW, ts_days_VIC, ts_days_SA) |>
  mutate(Date = as.Date(time),
         gauge = paste(site, gauge_type, sep = "_"))

#Wide format for EWR tool
hydrographs_LowerDarling_wide <- hydrographs_LowerDarling_long |>
  select(Date, value, gauge) |>
  pivot_wider(names_from = gauge, values_from = value)

write.csv(hydrographs_LowerDarling_wide, file.path(project_dir, "data", "hydrographs", paste0("hydrographs_LowerDarling_20250516", ".csv")), row.names = FALSE)

write.csv(get_ewr_table(), file.path(project_dir, "data", paste0("parameter_sheet_20250516", ".csv")), row.names = FALSE)

```

## Plot Hydrographs:
```{r}
hydrographs_LowerDarling_long <- hydrographs_LowerDarling_long |>
  separate(col = station_name, into = c("site_name", NA), sep = "Downstream", remove = FALSE)|>
  mutate(site_name = ifelse(is.na(site_short_name), site_name, site_short_name))|>
  mutate(site_name2 = ifelse(gauge_type == "flow", paste(site_name, gauge_type), site_name))


hydrographs <- filter(hydrographs_LowerDarling_long, Date > "2024-06-30")|>
ggplot(aes(x = Date, y = value, color = gauge_type)) +
  geom_line() +
  facet_wrap(site_name2 ~ ., scales = 'free_y', labeller = label_wrap_gen(10)) +
  theme_hydrobot()

ggsave(hydrographs, filename = file.path(project_dir, "data", "plots", paste0("hydrographs_LowerDarling_20250508", ".png")),
       width = 12, height = 12)

```


