---
title: "Simple toolkit template"
echo: true
---

## Getting started

There are many options for each of the steps here, this is a minimal, quick example designed to get you something that works quickly and can be modified. *In particular, we would expect the aggregation sequences and figures to need to be tailored for each analysis.*

See (in development) [documentation website](https://mdbauth.github.io/WERP_toolkit_demo/) for more detailed examples and discussion of capabilities.

Each step saves metadata about its parameters, which can be used for parameter-based runs to replicate, as well as variables to use in plots.

```{r setup}
#| include: false 
#| #| warning: false 
#| #| message: false
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file(), cache.lazy = FALSE, warning = FALSE, message = FALSE)

```

### Load libraries

```{r}
#| message: false
library(HydroBOT)
library(sf)
library(dplyr)
library(ggplot2)
library(tidyverse)
#devtools::install_github("galenholt/hydrogauge")
library(hydrogauge)
```


## User arguments

As a user, we need to set

-   Paths to input

-   Paths to save output

-   Module arguments (minimal)

-   Aggregation sequence and functions

-   Figures (typically interactive)

## Paths

```{r}
# Outer directory for project data
project_dir = getwd()

# Hydrographs (expected to exist already)
hydro_dir = file.path(project_dir, 'data', 'hydrographs')

# Generated data
EWR_subdir <- 'LowerDarling'

# EWR outputs (created here automatically in controller, need to read from here to aggregate)
ewr_results <- file.path(project_dir, 'data')

# outputs of aggregator. There may be multiple modules
agg_results <- file.path(project_dir, 'data', 'aggregator_output')
```

# EDIT CAUSAL NETWORKS
```{r}

# sdl_units_geodrop <- sdl_units |>
#   st_drop_geometry()
# 
# unique(sdl_units_geodrop$SWSDLName)
# unique(ewr2obj$LTWPShortName)
# unique(obj2target)

ewr2obj <- HydroBOT::causal_ewr$ewr2obj

#There are targets missing in here:
obj2target <- HydroBOT::causal_ewr$obj2target

#Check where env_obj is in ewr2obj but not obj2target, and vice versa
#all_links <- left_join(ewr2obj, obj2target, by = join_by(planning_unit_name, state, LTWPShortName, env_obj, Target, SWSDLName))

#THESE ARE A WORRY: i have not fixed this 08/05/25
#env_obj is in ewr2obj but not obj2target
unique(ewr2obj$env_obj)[!unique(ewr2obj$env_obj) %in% unique(obj2target$env_obj)]
# "NF3.NF4" "NF4.NF5" "NV2A"    "NV4"     "NV4E"    "OS3"     "NF"      "Q-NF20"  "Q-EF4"   "F4"      "O2"      "PEF 1a" 
# "PEF 1b"  "PEF 1c"  "PEF 3a"  "PEF 3b"  "PEF 4"   "PEF 5"   "O5"      "C1"      "F3"      "F5"      "F6"      "O1"     
# "O3"      "O4"      NA
# I did fix: "NF3.NF4" "NF4.NF5" "NV2A" and "NV4E" # these are just typos

#env_obj is in obj2target  but not ewr2obj
unique(obj2target$env_obj)[!unique(obj2target$env_obj) %in% unique(ewr2obj$env_obj)]
#"OS3a"   "Q-WB4"  "Q-NV3"  "PF 2.2" "PF 2.3" "PF 1.3"

#another small problem: i have not fixed this 08/05/25
ewr2obj |>
  filter(is.na(Target))
#This has no env_obj or target


# Some easy fixes:
#PROBLEM 1: "NF3.NF4", "NF4.NF5" in the "Murray Lower Darling" these need to be split. e.g: 
# temp <-  ewr2obj |>
#   filter(env_obj %in% c("NF3.NF4", "NF3", "NF4", "NF4.NF5", "NF4", "NF5"))|>
#   filter(LTWPShortName == "Murray Lower Darling") |>
#   group_by(ewr_code_timing, planning_unit_name, gauge)|>
#   reframe(n = n(), 
#           env_objs = paste(env_obj, collapse = "_"))

Fix1_ewr2obj <-  ewr2obj |>
  filter(env_obj %in% c("NF3.NF4", "NF4.NF5")) |> # only in the "Murray Lower Darling"
  separate(col = env_obj, into = paste0("env_obj", 1:2), sep = "[.]", extra = "merge")

Fix1_ewr2obj <- pivot_longer(Fix1_ewr2obj, cols = env_obj1:env_obj2, names_to = "delete", values_to = "env_obj", values_drop_na = TRUE)|>
  select(-delete)

# add in fix 1:
ewr2obj <- ewr2obj |>
  filter(!env_obj %in% c("NF3.NF4", "NF4.NF5"))|>
  bind_rows(Fix1_ewr2obj)

#Check for cases where an env_obj maps to different targets in different catchments - these need to be unique
temp <- obj2target |>
  distinct(env_obj, Target) |>
  group_by(env_obj) |>
  mutate(n = n()) |>
  filter(n > 1)

# Problem 2:
#"NV2A" and "NV4E" vs "NV2a" and "NV4e"
ewr2obj <- ewr2obj |>
  mutate(env_obj = case_when(env_obj == "NV2A" ~ "NV2a", 
                             env_obj ==  "NV4E" ~ "NV4e", 
                             .default = env_obj))

#sort(unique(ewr2obj$env_obj))

# Problem 3:
# THESE env_obj NEED TO BE UNIQUE:
# "P1" "F1" "1"  "2"  "3"  "4" 
#sort(unique(obj2target$env_obj)) #which are only numbers? (didnt get this to work: mutate(only_numbers = grepl("^[A-Za-z]+$", env_obj, perl = T)))
# These may as well be unique:
# "1","2","2.1","2.2","2.3","2.4","3","4","5","6"
# TOGETHER:
make_unique <- c("1","2","2.1","2.2","2.3","2.4","3","4","5","6","P1","F1")

ewr2obj <- ewr2obj |>
  mutate(Target_short = case_when(Target == "Ecosystem function" ~ "EF", 
                                  Target ==  "Native fish" ~ "NF", 
                                  Target ==  "Native vegetation" ~ "NV", 
                                  Target ==  "Waterbirds" ~ "WB", 
                                  Target ==  "Other species" ~ "OS"))|>
  mutate(env_obj_new = ifelse(env_obj %in% make_unique,
                          paste(substring(SWSDLName, 1, 1), "-", Target_short, env_obj, sep = ""), 
                          env_obj))

#NEW obj2target
#paste(names(obj2target), collapse = ", ")
#"planning_unit_name, LTWPShortName, env_obj, Specific_goal, Objective, Target, state, SWSDLName"
obj2target <- ewr2obj |>
  distinct(planning_unit_name, LTWPShortName, env_obj, Target, state, SWSDLName)

causal_ewr$ewr2obj <- ewr2obj
causal_ewr$obj2target <- obj2target

```

## Module controls

EWR currently, will add others as available. For this demo, we'll return only the summary and yearly outputs, though all EWR outputs are available.
The `returnType` argument keeps them in-memory rather than saving and reading back in. To demonstrate, I'm not doing that but it would make sense for small runs.

```{r}
outputType <- list('summary', 'yearly')
returnType <- list('none') # list('summary', 'yearly')
```

## Aggregation sequencing

-   Sequence of steps

-   Sequence of functions

There are *many* ways to specify, these are simple.

```{r}
sdl_units$sWSDLName <- sdl_units$SWSDLName
```


```{r}

aggseq_1 <- list(all_time = 'all_time',
               #ewr_code = c('ewr_code_timing', 'ewr_code'), commented out see if this works
                    #planning_unit = planning_units,
               sWSDLName = sdl_units,
                    
               env_obj =  c('ewr_code_timing', "env_obj"),
                    #Specific_goal = c('env_obj', "Specific_goal"),
               Target = c('env_obj', 'Target'),
               
               mdb = basin#,
               #target_5_year_2024 = c('Target', 'target_5_year_2024')
               )


funseq_1 <- list('ArithmeticMean', 
               #'ArithmeticMean',
                    #'ArithmeticMean',
               'ArithmeticMean', # Was 'SpatialWeightedMean' in gap
                    #'ArithmeticMean',
               'ArithmeticMean',
                    #'ArithmeticMean',
               'ArithmeticMean',
               
               'SpatialWeightedMean'#,
               #'ArithmeticMean'
               )

#Gauge scale
aggseq_2 <- list(all_time = 'all_time',
               #ewr_code = c('ewr_code_timing', 'ewr_code'),
               #planning_unit = planning_units,
               env_obj =  c('ewr_code_timing', "env_obj"),
               #sdl_units = sdl_units,
               #Specific_goal = c('env_obj', "Specific_goal"),
               Target = c('env_obj', 'Target')#,
               #mdb = basin,
               #target_5_year_2024 = c('Target', 'target_5_year_2024')
               )


funseq_2 <- list('ArithmeticMean', 
               #'ArithmeticMean',
               #'ArithmeticMean',
               'ArithmeticMean',
               #'SpatialWeightedMean',
               #'ArithmeticMean',
               'ArithmeticMean'#,
               #'SpatialWeightedMean',
               #'ArithmeticMean'
               )

aggseq_3 <- list(all_time = 'all_time')#,

funseq_3 <- list('ArithmeticMean')#, 

```

# Run the toolkit

Typically automatically

-   Run hydrographs through modules

-   Aggregate and save responses

-   Comparisons/analyses

## Run modules

-   Currently just EWR

-   Where are hydrographs

-   Where to save output

*Using `rparallel = TRUE` here would run in parallel over all scenarios.*
```{r}
#| message: false
future::plan(future::multisession)

ewr_out <- prep_run_save_ewrs(hydro_dir = file.path(hydro_dir), 
                              output_parent_dir = file.path(ewr_results), 
                              #output_subdir = EWR_subdir,
                              outputType = outputType,
                              returnType = returnType,
                              model_format = "Standard time-series",
                              rparallel = TRUE)

```

#have a look
```{r}
yearly <- read.csv(file.path(ewr_results, "module_output", "EWR", "hydrographs_LowerDarling_20250408", "yearly.csv"))

head(yearly)

```


## Aggregate

Here, unlike the EWR tool, we will just return these to the active R session instead of saving.

-   Sequence (levels of each theme, space, time axis)

-   Aggregation functions at each step

-   Produces a list, with each sheet a step in the aggregation

```{r}
agged_data_1 <- read_and_agg(datpath = file.path(project_dir, "data", "module_output", "EWR"),
           type = 'achievement',
           geopath = bom_basin_gauges,
           causalpath = causal_ewr,
           groupers = 'scenario',
           aggCols = 'ewr_achieved',
           group_until = list(SWSDLName = is_notpoint, planning_unit_name = is_notpoint, gauge = is_notpoint),
           pseudo_spatial = 'sdl_units',
           aggsequence = aggseq_1,
           funsequence = funseq_1,
           saveintermediate = TRUE,
           namehistory = FALSE,
           keepAllPolys = FALSE,
           returnList = TRUE,
           savepath = file.path(agg_results, "sdl_target"))

agged_data_2 <- read_and_agg(datpath =  file.path(project_dir, "data", "module_output", "EWR"), 
           type = 'achievement',
           geopath = bom_basin_gauges,
           causalpath = causal_ewr,
           groupers = 'scenario',
           aggCols = 'ewr_achieved',
           group_until = list(SWSDLName = is_notpoint, planning_unit_name = is_notpoint, gauge = is_notpoint),
           pseudo_spatial = 'sdl_units',
           aggsequence = aggseq_2,
           funsequence = funseq_2,
           saveintermediate = TRUE,
           namehistory = FALSE,
           keepAllPolys = FALSE,
           returnList = TRUE,
           savepath = file.path(agg_results, "gauge_target"))

agged_data_3 <- read_and_agg(datpath =  file.path(project_dir, "data", "module_output", "EWR", "hydrographs_LowerDarling_20250408"), 
           type = 'achievement',
           geopath = bom_basin_gauges,
           causalpath = causal_ewr,
           groupers = 'scenario',
           aggCols = 'ewr_achieved',
           #group_until = list(SWSDLName = is_notpoint, planning_unit_name = is_notpoint, gauge = is_notpoint),
           #pseudo_spatial = 'sdl_units',
           aggsequence = aggseq_3,
           funsequence = funseq_3,
           saveintermediate = TRUE,
           namehistory = FALSE,
           keepAllPolys = FALSE,
           returnList = TRUE,
           savepath = file.path(agg_results, "gauge_all_time"))
```

The new scenario names ensure uniqueness, but can be ugly. This could be avoided with a named list to the `scenarios` argument in `prep_run_save_ewrs`, but a robust, general approach is a work in progress.

```{r}
# The new scenario names ensure uniqueness, but can be ugly
agged_data <- purrr::map(agged_data, 
                         \(x) x |> 
                           mutate(scenario = stringr::str_remove_all(scenario, "_[A-z]*[0-9]*")))
```

# Outputs

-   Standardised formats using `plot_outcomes` function

-   Different plots for different purposes

-   Small selection follows

*These are simply here as examples to make sure the toolkit works and illustrate some options. They are not a definitive set, and the figures created for any analysis should be tailored for that analysis.*

We can extract scenario information from the metadata to use when plotting.

And we also set some standard scenario ordering and color palettes.

```{r}
#| message: false
# Scenario data
scenarios <- jsonlite::read_json(file.path(hydro_dir, 
                                           'scenario_metadata.json')) |> 
  tibble::as_tibble() |> 
  tidyr::unnest(cols = everything()) |> 
  bind_rows(tibble::tibble(scenario_name = "MAX", flow_multiplier = Inf))

sceneorder <- forcats::fct_reorder(scenarios$scenario_name, scenarios$flow_multiplier)

scene_pal <- make_pal(unique(scenarios$scenario_name), 
                      palette = 'ggsci::nrc_npg', 
                      refvals = 'base', refcols = 'black', includeRef = TRUE)

```

This is just for looks- some of the target names are incredibly long, make them shorter.
And make a grouping variable for the environmental objectives. This is nearly the same as Target, but we want to have the individual env_obj too.

```{r}
# make the super long names shorter but less useful.
basin_to_plot <- agged_data$mdb %>% 
  dplyr::filter(!is.na(Objective)) %>% 
  dplyr::mutate(Objective = stringr::str_trunc(Objective, 15)) %>% 
  dplyr::group_by(scenario, Objective) %>% 
  dplyr::mutate(id = as.character(row_number())) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(Objective = stringr::str_c(Objective, '_', id)) %>% 
  dplyr::select(-id)

# Create a grouping variable
obj_sdl_to_plot <- agged_data$sdl_units |>
  left_join(scenarios, by = c('scenario' = 'scenario_name')) |>
  dplyr::mutate(env_group = stringr::str_extract(env_obj, '^[A-Z]+')) |>
  dplyr::filter(!is.na(env_group)) |>
  dplyr::arrange(env_group, env_obj)
```

## Input hydrographs

```{r}
#| echo: false
scenehydros <- read_hydro(hydro_dir, long = TRUE, format = 'csv')|> 
                           mutate(scenario = stringr::str_remove_all(scenario,
                                                                     "_[A-z]*[0-9]*"))

hydros <- scenehydros |> 
  plot_outcomes(outcome_col = 'flow',
                x_col = 'Date',
                colorset = 'scenario',
                pal_list = scene_pal,
                facet_wrapper = 'gauge')

hydros
```

## Maps and spatial scaling

```{r}
#| message: false

# Quick plot
map_example <- agged_data$sdl_units |>
  dplyr::filter(env_obj == 'NF1') |> # Need to reduce dimensionality
  plot_outcomes(outcome_col = 'ewr_achieved',
                plot_type = 'map',
                colorset = 'ewr_achieved',
                pal_list = list('scico::berlin'),
                facet_col = 'scenario',
                facet_row = 'env_obj',
                sceneorder = sceneorder,
                underlay_list = 'basin',
                overlay_list = list(overlay = dplyr::filter(
                  agged_data$env_obj,
                    env_obj == 'NF1'
                  ),
                  pal_list = 'scico::oslo',
                  outcome_col = 'ewr_achieved',
                  clip = TRUE)) +
  ggplot2::theme(legend.position = 'bottom')

map_example
```

## Bars- SDL units and scenarios

SDL unit differences in all environmental objectives

```{r}
#| echo: false
catchcompare <- agged_data$sdl_units %>% 
  group_by(SWSDLName, scenario) %>% 
  summarise(ewr_achieved = sum(ewr_achieved, na.rm = TRUE)) %>% 
  ungroup() %>% 
plot_outcomes(outcome_col = 'ewr_achieved',
              colorset = 'SWSDLName',
              pal_list = list("calecopal::lake"),
              sceneorder = sceneorder,
              position = 'dodge')

catchcompare
```

## Objective and scenario comparisons

::: panel-tabset
### Basin

```{r}
#| echo: false
objective_comp <- basin_to_plot |> 
  dplyr::filter(scenario != 'MAX') |> # max scenario is just all 1, so boring.
  plot_outcomes(
                 outcome_col = 'ewr_achieved',
                 x_col = 'Objective',
                 colorset = 'scenario',
                 pal_list = scene_pal,
                 sceneorder = sceneorder,
              position = 'dodge'
              )

objective_comp
```

### SDL units

```{r}
#| echo: false
#| message: false
# Create a palette list
env_pals = list(EB = 'grDevices::Grays',
                EF = 'grDevices::Purp',
                NF = 'grDevices::Mint',
                NV = 'grDevices::Burg',
                OS = 'grDevices::Blues',
                WB = 'grDevices::Peach')

# need to facet by space sdl unit and give it the colorgroup argument to take multiple palettes
sdl_stack <- obj_sdl_to_plot |>
  plot_outcomes(outcome_col = 'ewr_achieved',
                colorgroups = 'env_group',
                colorset = 'env_obj',
                pal_list = env_pals,
                facet_col = 'SWSDLName',
                facet_row = '.',
                sceneorder = sceneorder)
sdl_stack +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 0.5))
```
:::

## Lines and baseline

Change relative to baseline available to all plots

-   Disproportionate response

```{r}
#| echo: false

line_pals = list(EB = 'scico::lisbon',
                EF = 'scico::acton',
                NF = 'scico::bamako',
                NV = 'scico::batlow',
                OS = 'scico::lajolla',
                WB = 'scico::hawaii')


sdl_line_groups_facet <- obj_sdl_to_plot |>
  plot_outcomes(outcome_col = 'ewr_achieved',
                x_col = 'flow_multiplier',
                outcome_lab = 'Proportion met',
                x_lab = 'Change in flow',
                transx = 'log10',
                transy = 'log10',
                color_lab = 'Environmental\ngroup',
                colorgroup = 'env_group',
                colorset = 'env_obj',
                pal_list = line_pals,
                facet_row = 'SWSDLName',
                facet_col = 'env_group',
                base_list = list(base_lev = 'base',
                                 comp_fun = 'relative',
                                 add_eps = min(obj_sdl_to_plot$ewr_achieved[obj_sdl_to_plot$ewr_achieved > 0], 
                              na.rm = TRUE)/2,
                              group_cols = c('env_obj', 'polyID'))
  )

sdl_line_groups_facet + theme(legend.position = 'none') + 
  coord_cartesian(ylim = c(0, 10))
```

## Fits

Smoothed fit of all environmental objectives in each group

-   Relative to baseline

```{r}
#| echo: false
#| message: false
sdl_fit_sdl <- obj_sdl_to_plot |>
  plot_outcomes(outcome_col = 'ewr_achieved',
                x_col = 'flow_multiplier',
                outcome_lab = 'Proportion met',
                x_lab = 'Change in flow',
                transx = 'log10',
                transy = 'log10',
                color_lab = 'SDL unit',
                colorset = 'SWSDLName',
                pal_list = list("ggsci::default_jama"),
                facet_wrapper = 'env_group',
                base_list = list(base_lev = 'base',
                                 comp_fun = 'relative',
                                 add_eps = min(obj_sdl_to_plot$ewr_achieved[obj_sdl_to_plot$ewr_achieved > 0], 
                              na.rm = TRUE)/2,
                              group_cols = c('env_obj', 'polyID')),
                smooth_arglist = list())

suppressWarnings(print(sdl_fit_sdl + 
  coord_cartesian(ylim = c(0, 10))))
```

## Causal networks

This is not quite as integrated into the standard plotting as it could be. We need to do a bit of data wrangling, but not too much.

First, we build the networks, we need to not double up when aggregation is spatial.

```{r}
# Which steps are theme? This glosses over potential changes with space
stepdim <- identify_dimension(aggseq, causal_ewr)

themesteps <- stepdim == 'theme'
names(themesteps) <- names(aggseq)

themeseq <- aggseq[themesteps]

ewr_edges <- make_edges(dflist = causal_ewr, 
                        fromtos = themeseq[2:length(themeseq)],
                        gaugefilter = '421001')

nodes <- make_nodes(ewr_edges)
```

Then we add the outcome values to the nodes so we can use them as colour. 

```{r}
#| message: false
#| warning: false

# What is the column that defines the value?
valcol <- 'ewr_achieved'

# Get the values for each node
targetlevels <- names(themesteps)[themesteps]

# If there were more than 1 aggfun per step, we'd need to choose a single path through
# themefuns <- funseq[themesteps]

aggvals <- extract_vals_causal(agged_data, funseq, valcol, 
                               targetlevels = targetlevels)

# cut to relevant gauge (or no gauge for higher spatial levels)
  # THis is slow- I think what we'd actually do is find the matching sdl and cut to that earlier
gaugematch <- st_intersects(bom_basin_gauges[bom_basin_gauges$gauge == '421001',],
                            aggvals, sparse = FALSE)

aggvals <- aggvals[as.vector(gaugematch),] |>
  st_drop_geometry()
# aggvals <- aggvals %>% dplyr::filter(gauge == '421001' | is.na(gauge)) %>% 
#   dplyr::select(-gauge)

# join to the nodes
nodes_with_vals <- nodes |> 
  dplyr::filter(NodeType != 'ewr_code_timing') |> 
  dplyr::left_join(aggvals) |>
  dplyr::filter(!is.na(scenario))
```

::: panel-tabset

### Down 4

```{r}
#| echo: false
aggNetworkdown <- make_causal_plot(nodes = dplyr::filter(nodes_with_vals, 
                                        scenario == 'down4'),
                 edges = ewr_edges,
                 edge_pal = 'black',
                 node_pal = list(value = 'scico::tokyo'),
                 node_colorset = 'ewr_achieved',
                 render = FALSE)

DiagrammeR::render_graph(aggNetworkdown)
```

### Baseline

```{r}
#| echo: false
aggNetworkbase <- make_causal_plot(nodes = dplyr::filter(nodes_with_vals, 
                                        scenario == 'base'),
                 edges = ewr_edges,
                 edge_pal = 'black',
                 node_pal = list(value = 'scico::tokyo'),
                 node_colorset = 'ewr_achieved',
                 render = FALSE)

DiagrammeR::render_graph(aggNetworkbase)
```

### Up 4

```{r}
#| echo: false
aggNetworkup <- make_causal_plot(nodes = dplyr::filter(nodes_with_vals, 
                                        scenario == 'up4'),
                 edges = ewr_edges,
                 edge_pal = 'black',
                 node_pal = list(value = 'scico::tokyo'),
                 node_colorset = 'ewr_achieved',
                 render = FALSE)

DiagrammeR::render_graph(aggNetworkup)
```
:::

